---
title: "WUP_04"
author: "Tyler Frasier "
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
gss=read.csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  

$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The responsible variable is "WORDSUM" which is numerical. Thhe predictopr variable is "DEGREE" which is categorical. therefore i will use numerical~cat. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->



```{r}
ggplot(gss, aes(x = DEGREE, y = WORDSUM)) +
 geom_boxplot(fill = "lightblue") +
 xlab("DEGREE") +
 ylab("WISDUM") +
 ggtitle("Box PLot of WORDSUM by DEGREE")

```

Thhis plot shows based on your degree level it shows average scores. bachelor, graduate, and junior college have the highest scores. they average the same, LT high school and high school hhave the lowest scores. they average about the same.

### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE, data = gss)
```

This table shows thhe difference degree's median and mean scores. Bachelor and junior college all have roughly the same average mean and median scores. they average about 9. The graduates have the highest score which is 10. while high school and LT high school have roughly the same scores. Thhey average 8.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data=gss)
summary.aov(model1)
```

```{r}
plot(TukeyHSD(model1))
print(TukeyHSD(model1))
```


This graph shows the p value whichh is 0.000236. Thherefore i will reject the null hypothesis. Because it is lower than 0.05. There is a difference, the only place is lths to g and hhs to g. 

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
The original question is, Is thhere a differencd in the average scores on th WORDSUM test based on level of schooling degree. There is a difference in average score on thhe WORDSUM test between the different schhooling degrees. there is a big difference in hs to g and Lths to scores. Since the p value is 0.000236, whichhh is lower than 0.05. i will reject the null hypothesis.